### 2015-06-16 reports/paper/conclusion.tex
paper update

\chapter{Conclusion}
In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

This system can be used in HCI applications that incorporate sound position and movement information, such as drawing with music, or can be used to create AI games with physical pieces where the computer controls some of the pieces. This system can also be used to in Augmented Reality (AR) applications that allow user computer interaction with music tags.

For future directions, this system can be extended with another microphone on each array to perform 3D localization. Another possible direction is to investigate multi-source localization using digitally tagged audio sources.

>>>>

\chapter{Conclusion}
In this thesis, we described and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and showed that the localization accuracy varies with the array geometry and the location of the sound source (both the distance and the angle). In particular, regions close to the line connecting two microphones are generally more difficult to localize than regions close the line bisecting two microphones. This problem can be alleviated by employing multiple microphones that are placed at large distance from each other, but such arrangement affects the portability of the system. As a trade off between accuracy and portability, we demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region. Since two arrays generate their likelihood maps individually, this system does not require synchronized clock between the arrays, which made the overall system easier to design and more portable.

This system can be used in HCI applications that incorporate sound position and movement information. One can use this system to build virtual drawing applications where the user can draw with music, without physically touching the computer. One can also use this system to design interactive AI games. For example, a chess game with physical pieces can be developed where each piece is equipped with a motor and a music tag. Since the computer knows where all the pieces are, it knows which piece the user has moved and can make its corresponding move. A similar example is AI toy car racing game, where the player controls one car and the computer controls another car. With real time location information, the computer can control one car to compete with the player on a racing track. This system can also be used to in Augmented Reality (AR) applications. For example, users can wear a music tag on the finger, and then the system would be able to track finger movement. It can be used as virtual mouse, and applications can be developed to detect finger gesture.

For future directions, one can extend the system with another microphone on each array to perform 3D localization. However, the number of grid points increases exponentially with the number of dimension, which means the calculation time also increases exponentially. Therefore, it is important to research more efficient ways to find the intersection of hyperbolic cones the in 3D space.  Since the likelihood calculation on each grid point is independent of each other, one straightforward way to speedup calculation is to employ hardware accelerated parallel computing technologies such as GPU or FPGA. Another useful direction is to investigate multi-source localization using digitally tagged audio sources.

### 2015-06-16 reports/paper/experiment.tex
paper update

\chapter{Experiment}
We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.

\section{Setup}
\subsection{Point localization}

>>>>

\chapter{Experiment}
\section{Setup}
\input{./system}

We conducted two sets of experiments to evaluate the system's localization accuracy: one on point localization and the other on movement tracking.

\subsection{Point localization}

### 2015-06-16 reports/paper/prior.tex
paper update

Let $s(t)$ denote the audio signal. The received signal $\bar s(t) $ at microphone $i$ can be modeled as:
\[
\bar s_i(t) = \frac{s(t)}{d_i} + \delta_i(t)
\]
where $d_i$ is the distance between the audio source and microphone $i$, and $\delta_i(t)$ represents Gaussian noise.
The energy received at microphone $i$ is:
\begin{eqnarray}
E_i & = & \int \bar s_i(t)^2 dt \\
\label{eqn:ild1}
& = & \frac{1}{d_i^2} E_s + E_{\sigma_i}
\end{eqnarray}
It shows that the energy decays with the square of the distance. With two microphones, not considering the noise $\delta_i(t)$, energy received at both microphones has to satisfy equation~\ref{eqn:ild1}:
\begin{eqnarray}
E_1d_1^2 &=& E_2d_2^2\\
\label{eqn:ild}
\frac{E_1}{E_2} &=& \frac{d_2^2}{d_1^2}
\end{eqnarray}
It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $E_1 \ne E_2$ and form a line when $E_1 = E_2$~\cite{ild:1}. With two microphones all points on this curve generate the same energy ratio and can not be distinguished from each other. Multiple approaches have been investigated to eliminate this ambiguity. \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location. \cite{ild:2} combined ILD with Interaural Time Difference to estimate the source direction (i.e azimuth).

ILD approach relies on the accurate measurement of the received energy ratio between microphone pairs. Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured energy ratio. We find this approach to be too restrictive for our system, since in an interactive system we can not control whether or not the user places any obstacle between the sound source and the microphones.

@@

To see the reasoning behind different prefiltering approaches, we separate the magnitude part from the phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
\begin{eqnarray}
\label{eqn:phat1}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) |X_1(\omega)||X_2(\omega)| e^{j(\omega\tau - (\angle{X_2(\omega)} - \angle{X_1(\omega)}) } d\omega\\
&=& \int_{-\infty}^\infty \underbrace{W(w)|X_1(\omega)| |X_2(\omega)|}_{\mbox{weighting}} \underbrace{\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))}_{\mbox{phase error}} d\omega
\end{eqnarray}
We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals. When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, $\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega)) = 0$, and $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) = 1$. When $\dot \tau$ differs from the true arrival time difference, $\cos(\omega \dot \tau - (\angle X_2(\omega) - \angle X_1(\omega))) < 1$. Therefore, $\cos(\omega\tau - (\angle X_2(\omega) - \angle X_1(\omega)))$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency. The TDOA estimator essentially sums the weighted phase error at each frequency.

>>>>

For a point sound source in a direct field, the signal intensity decays proportional to the square of the distance between the source and the microphone. Let $I_i$ denote the received signal intensity at microphone $i$: 
\[
I_i \propto \frac{1}{d_i^2} I_s
\]
where $d_i$ is the distance between the audio source and microphone $i$, and $I_s$ is the sound intensity at the source.
With two microphones, the signal intensity received at both microphones has to satisfy equation~\ref{eqn:ild1}:
\begin{eqnarray}
I_1d_1^2 &=& I_2d_2^2\\
\label{eqn:ild}
\frac{I_1}{I_2} &=& \frac{d_2^2}{d_1^2}
\end{eqnarray}
It can be shown that, on a 2D plane, points satisfy equation~\ref{eqn:ild} form a circle when $I_1 \ne I_2$ and form a line when $I_1 = I_2$~\cite{ild:1}. With two microphones all points on this curve generate the same intensity ratio and can not be distinguished from each other. Multiple approaches have been investigated to eliminate this ambiguity. \cite{ild:1} employed multiple microphones and used the intersection of circles from each microphone pair to estimate the source location. \cite{ild:3} combined ILD with Interaural Time Difference (ITD) to estimate the source direction (i.e azimuth). Instead of solving the intersection of circles, \cite{ild:4} employed machine learning techniques to automatically learn the mapping from ILD and ITD features to location coordinate. It uses four microphones and requires a training phase, during which the sound source is manually placed at predetermined locations for the system to learn the parameters that map from feature space to the sound location.

ILD approach relies on the accurate measurement of the received intensity ratio between microphone pairs. Any obstacle object between the sound source and any microphone would produce a significant distortion in the measured intensity ratio. We find this approach to be too restrictive for our system, since in an interactive system we can not control whether or not the user places any obstacle between the sound source and the microphones.

@@

To see the reasoning behind different prefiltering approaches, we separate the magnitude part from the phase part of $X_1(\omega)$ and $X_2(\omega)$ in Equation~\ref{eq:gcc2}:
\begin{eqnarray}
\label{eqn:phat1}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) |X_1(\omega)||X_2(\omega)| e^{j(\omega\tau - (\angle{X_2(\omega)} - \angle{X_1(\omega)}) } d\omega\\
&=& \int_{-\infty}^\infty \underbrace{W(w)|X_1(\omega)| |X_2(\omega)|}_{\mbox{weighting}} \cos(\Theta_\epsilon) d\omega
\end{eqnarray}
Where $\Theta_{e}$ is the phase error:
\[
\Theta_\epsilon = \omega\tau - (\angle X_2(\omega) - \angle X_1(\omega))
\]
We can look at the real part of equation~\ref{eqn:phat1} only since both $x_1(t)$ and $x_2(t)$ are real valued signals. When $\dot \tau$ is the true arrival time difference between $x_1(t)$ and $x_2(t)$, phase error $\Theta_\epsilon  = 0$, and $\cos(\Theta_\epsilon) = 1$. When $\dot \tau$ differs from the true arrival time difference, $\cos(\Theta_\epsilon) < 1$. Therefore, $\cos(\Theta_\epsilon)$ can be seen as a measure of the phase error, and $W(w)|X_1(\omega)||X_2(\omega)|$ describes how the error should be weighted at each frequency. The TDOA estimator essentially sums the weighted phase error at each frequency.
